---
title: "Confederate Statues in the US"
editor_options: 
  chunk_output_type: console
output: html_document
---

```{r prep, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  tidy = FALSE, results = "asis"
)

library(tidyverse)
```

This is so wild. Let this sink in.

This was motivated by my recent trip to South Carolina and seeing my first confederate statue. I am not sure why we glorify these people who fought for people to be enslaved.

Check these out btw:

1. https://www.youtube.com/watch?v=dOkFXPblLpU&feature=emb_title
2. https://www.youtube.com/watch?v=WClgR6Q0aPE
3. https://www.theguardian.com/us-news/2017/aug/16/confederate-monuments-civil-war-history-trump

(1) is about how the United Daughters of the Confederacy brainwashed the youth by pretending slavery wasn't the cause of the Civil War
(2) talks about the reason why each peek of confederate statues were created
(3) also talks about each peek and why these statues were created

```{r}
all_dates = read.csv("data/confederate_statue_dates.csv") %>% as_tibble()
```

```{r graph1, fig.width=20, fig.height=15}
all_dates %>%
  ggplot(aes(x = value3)) +
  geom_histogram(binwidth = 5, fill = "red") +
  scale_x_continuous(breaks = seq(min(all_dates$value3), max(all_dates$value3), by = 5)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size=25, angle = 90),
        axis.text.y = element_text(size=30),
        title = element_text(size=30)) +
  ylab("# of Statues") +
  ggtitle("Number of Confederate Statues Over Time")
```

It's just something to think about...

Also web scraping is fun


