---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# When were Confederate Statues made?

<!-- badges: start -->
[![R tests](https://github.com/cavandonohoe/confederate_statues/actions/workflows/test.yml/badge.svg)](https://github.com/cavandonohoe/confederate_statues/actions/workflows/test.yml)
[![Refresh dataset](https://github.com/cavandonohoe/confederate_statues/actions/workflows/refresh-data.yml/badge.svg)](https://github.com/cavandonohoe/confederate_statues/actions/workflows/refresh-data.yml)
<!-- badges: end -->

A small R pipeline that scrapes Wikipedia's lists of Confederate monuments and memorials, extracts the year each one was installed, and plots the timeline.


```{r prep, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  tidy = FALSE, results = "asis"
)

library(tidyverse)
```

This is so wild. Let this sink in.

This was motivated by my recent trip to South Carolina and seeing my first confederate statue. I am not sure why we glorify these people who fought for people to be enslaved.

Check these out btw:

1. https://www.youtube.com/watch?v=dOkFXPblLpU
2. https://www.youtube.com/watch?v=WClgR6Q0aPE
3. https://www.theguardian.com/us-news/2017/aug/16/confederate-monuments-civil-war-history-trump

((1)) is about how the United Daughters of the Confederacy brainwashed the youth by pretending slavery wasn't the cause of the Civil War

((2)) talks about the reason why each peek of confederate statues were created

((3)) also talks about each peek and why these statues were created

```{r}
all_dates = read.csv("data/confederate_statue_dates.csv") %>% as_tibble()

max_statues = all_dates %>% group_by(year) %>%
  count %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  filter(n == max(n))

mode = max_statues %>% 
  pull(year)

with_max = max_statues %>% 
  pull(n)
```

Also, we witness the most statues being built in a single year in `r mode` with `r with_max` statues.

These bins are of width = 5 years.

```{r graph1, fig.width=20, fig.height=15}
all_dates %>%
  ggplot(aes(x = year)) +
  geom_histogram(binwidth = 5, fill = "red") +
  scale_x_continuous(breaks = seq(min(all_dates$year), max(all_dates$year), by = 5)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size=25, angle = 90),
        axis.text.y = element_text(size=30),
        title = element_text(size=30)) +
  ylab("# of Statues") +
  ggtitle("Number of Confederate Statues Over Time")
```

It's just something to think about...

Also web scraping is fun.

## How the pipeline works

Pure functions live in `R/scrape.R` and are unit-tested against fixture HTML in `tests/testthat/` (no network calls during tests).

| Layer | File | What it does |
|---|---|---|
| Data layer | `R/scrape.R` | `parse_dates_from_text()` (pure), `grab_dates(url, cache_dir)` (network + parse), `scrape_all(urls)` (orchestrates all sources). Source URLs are the `STATE_URLS` constant. |
| Refresh | `scripts/refresh_data.R` | Re-runs `scrape_all()` and writes `data/confederate_statue_dates.csv`. Run with `--cache` to also save raw HTML under `data/raw_html/`. |
| Tests | `tests/testthat/test-scrape.R` | 9 cases covering year extraction, multi-year entries, citation-noise filtering, cache-dir behavior, and the URL constants. |
| Automation | `.github/workflows/test.yml` | Runs tests on push and PR. |
| Automation | `.github/workflows/refresh-data.yml` | Monthly cron + manual trigger that re-runs the refresh script and opens a PR if the CSV changed. |

To refresh locally:

```bash
Rscript scripts/refresh_data.R
```

To run the tests:

```r
testthat::test_dir("tests/testthat")
```

## Sources

- [Alabama](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Alabama)
- [Georgia](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Georgia)
- [Mississippi](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Mississippi)
- [North Carolina](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_North_Carolina)
- [South Carolina](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_South_Carolina)
- [Other states](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials)

