---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# When were Confederate Statues made?

<!-- badges: start -->
[![R tests](https://github.com/cavandonohoe/confederate_statues/actions/workflows/test.yml/badge.svg)](https://github.com/cavandonohoe/confederate_statues/actions/workflows/test.yml)
[![Refresh dataset](https://github.com/cavandonohoe/confederate_statues/actions/workflows/refresh-data.yml/badge.svg)](https://github.com/cavandonohoe/confederate_statues/actions/workflows/refresh-data.yml)
<!-- badges: end -->

A small R pipeline that scrapes Wikipedia's lists of Confederate monuments and memorials, extracts the year each one was installed, and plots the timeline.


```{r prep, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  tidy = FALSE, results = "asis"
)

library(tidyverse)
```

This is so wild. Let this sink in.

This was motivated by my recent trip to South Carolina and seeing my first confederate statue. I am not sure why we glorify these people who fought for people to be enslaved.

Check these out btw:

1. https://www.youtube.com/watch?v=dOkFXPblLpU
2. https://www.youtube.com/watch?v=WClgR6Q0aPE
3. https://www.theguardian.com/us-news/2017/aug/16/confederate-monuments-civil-war-history-trump

((1)) is about how the United Daughters of the Confederacy brainwashed the youth by pretending slavery wasn't the cause of the Civil War

((2)) talks about the reason why each peek of confederate statues were created

((3)) also talks about each peek and why these statues were created

```{r}
all_dates = read.csv("data/confederate_statue_dates.csv") %>% as_tibble()

max_statues = all_dates %>% group_by(year) %>%
  count %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  filter(n == max(n))

mode = max_statues %>% 
  pull(year)

with_max = max_statues %>% 
  pull(n)
```

Also, we witness the most statues being built in a single year in `r mode` with `r with_max` statues.

These bins are of width = 5 years.

```{r graph1, fig.width=20, fig.height=15}
all_dates %>%
  ggplot(aes(x = year)) +
  geom_histogram(binwidth = 5, fill = "red") +
  scale_x_continuous(breaks = seq(min(all_dates$year), max(all_dates$year), by = 5)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size=25, angle = 90),
        axis.text.y = element_text(size=30),
        title = element_text(size=30)) +
  ylab("# of Statues") +
  ggtitle("Number of Confederate Statues Over Time")
```

It's just something to think about...

Also web scraping is fun.

## How the pipeline works

The pipeline is split into four layers so each piece is independently
testable, and so a fresh checkout can regenerate the dataset offline
without hitting Wikipedia. Pure functions live in `R/` and are
unit-tested against fixture HTML in `tests/testthat/` (no network calls
during tests).

```
data-raw/*.html  ->  scrape_all()  ->  tidy_statues()  ->  data/confederate_statue_dates.csv
   (snapshots)    (R/scrape.R)       (R/clean.R)        (canonical dataset)
```

| Layer | File | What it does |
|---|---|---|
| Sources (raw) | `data-raw/*.html` | Checked-in HTML snapshots of each Wikipedia source page. Makes the pipeline reproducible offline. |
| Scrape | `R/scrape.R` | `parse_dates_from_text()` (pure), `grab_dates(url)` (parse single page, works on local files or URLs), `scrape_all(urls)` (orchestrates all sources). |
| Clean | `R/clean.R` | `tidy_statues(raw)` locks in the canonical schema: drops year outliers and duplicates, coerces types, sorts deterministically. |
| Output | `data/confederate_statue_dates.csv` | Cleaned dataset, regenerated from `data-raw/`. |
| Tests | `tests/testthat/test-scrape.R` | Parser cases: year extraction, multi-year entries, citation-noise filtering, cache-dir, URL constants. |
| Tests | `tests/testthat/test-clean.R` | Schema lock-in: column names, types, year-window filter, deduplication, whitespace trimming, sort order. |
| Refresh script | `scripts/refresh_data.R` | Parses `data-raw/*.html` -> cleans -> writes the CSV. Offline. |
| Refetch script | `scripts/refetch_raw.R` | Re-downloads HTML from Wikipedia into `data-raw/`. Network. |
| CI | `.github/workflows/test.yml` | Runs both test files on push and PR. |
| CI | `.github/workflows/refresh-data.yml` | Monthly cron: runs refetch -> refresh, opens a PR if anything changed. |

To regenerate the dataset locally (offline, from the checked-in HTML):

```bash
Rscript scripts/refresh_data.R
```

To also re-download fresh HTML from Wikipedia first:

```bash
Rscript scripts/refetch_raw.R && Rscript scripts/refresh_data.R
```

To run the tests:

```r
testthat::test_dir("tests/testthat")
```

## Limitations

A few things to keep in mind when reading the chart:

- **Wikipedia is the only source.** Coverage is uneven across states.
  Some states (Alabama, Georgia, Mississippi, the Carolinas) have
  dedicated list pages; everything else is rolled up into a single
  catch-all page that's far less complete.
- **Years come from a `(YYYY)` regex.** The parser extracts every
  4-digit token in parentheses from each entry line. Entries that
  store the year differently (a date range, a prose description,
  a footnote-only year) are silently missed. Entries that reference
  multiple years (installed in YYYY, replaced in YYYY) produce
  multiple rows.
- **Borderline pre-war entries are kept.** A handful of antebellum
  counties were later co-opted by Confederate naming (e.g. "Levy
  County (1845)"). `tidy_statues()` keeps anything from 1840 onward;
  this is a deliberate choice over a stricter Civil-War-onward cutoff,
  but it does mean a few rows aren't strictly "monuments built during
  or after the Confederacy".
- **No deduplication across source pages.** If the same monument
  appears on both a state-specific page and the catch-all "other"
  page, it'll show up twice. `tidy_statues()` only dedups within a
  given source.
- **The dataset reflects whatever Wikipedia looked like on the last
  refresh.** The monthly scheduled job in
  `.github/workflows/refresh-data.yml` keeps `data-raw/` and the CSV
  in sync, and the diff lands as a PR for review.

## REST API service

The cleaned dataset is also exposed by a small Python (FastAPI) REST
service that lives in `service/`. It demonstrates the full
data-engineering loop end to end: scrape (R) -> clean (R) -> CSV ->
ingest (Python) -> Postgres -> HTTP API.

| Endpoint              | What it returns                                |
|-----------------------|------------------------------------------------|
| `GET /statues`        | Paginated, filterable list (state, year, text) |
| `GET /statues/{id}`   | A single entry by content-hash id              |
| `GET /stats/by-decade`| Counts bucketed by decade                      |
| `GET /stats/by-state` | Counts grouped by source state                 |
| `GET /health`         | Liveness + DB connectivity probe               |
| `GET /metrics`        | Prometheus text exposition                     |
| `GET /docs`           | Swagger UI (interactive)                       |
| `GET /openapi.json`   | OpenAPI 3.1 schema for codegen                 |

Run it locally with Docker:

```bash
cd service
docker compose up --build
docker compose exec api statue-api ingest --csv /data/confederate_statue_dates.csv
open http://localhost:8000/docs
```

Then query it from anywhere. A few examples:

```bash
# Counts per decade -- one line per histogram bar
curl -s http://localhost:8000/stats/by-decade

# Filtered list: Georgia entries from the 1900-1910 spike
curl -s 'http://localhost:8000/statues?source=georgia&year_min=1900&year_max=1910&limit=10'

# Free-text search
curl -s 'http://localhost:8000/statues?q=Robert+E.+Lee&limit=5'
```

```python
import httpx, pandas as pd
df = pd.DataFrame(
    httpx.get("http://localhost:8000/statues", params={"limit": 500}).json()["items"]
)
df.groupby("source").size().sort_values(ascending=False)
```

```r
library(httr2)
decades <- request("http://localhost:8000") |>
  req_url_path("/stats/by-decade") |>
  req_perform() |>
  resp_body_json(simplifyVector = TRUE)
ggplot2::qplot(decades$items$decade_start, decades$items$count, geom = "col")
```

Stack: FastAPI, Pydantic, SQLAlchemy 2.0, Alembic, Postgres, structlog,
Prometheus, pytest, ruff, mypy, Docker, Fly.io. See
[`service/README.md`](service/README.md) for the full endpoint
reference (every query param, response shape, error case), more
client examples, the architecture diagram, configuration, and deploy
instructions.

## Sources

- [Alabama](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Alabama)
- [Georgia](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Georgia)
- [Mississippi](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Mississippi)
- [North Carolina](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_North_Carolina)
- [South Carolina](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_South_Carolina)
- [Other states](https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials)

