---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# When were Confederate Statues made?

<!-- badges: start -->
<!-- badges: end -->

R Project for web scraping Wikipedia looking for dates when confederate statues were erected


```{r prep, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  tidy = FALSE, results = "asis"
)

library(tidyverse)
```

This is so wild. Let this sink in.

This was motivated by my recent trip to South Carolina and seeing my first confederate statue. I am not sure why we glorify these people who fought for people to be enslaved.

Check these out btw:

1. https://www.youtube.com/watch?v=dOkFXPblLpU
2. https://www.youtube.com/watch?v=WClgR6Q0aPE
3. https://www.theguardian.com/us-news/2017/aug/16/confederate-monuments-civil-war-history-trump

((1)) is about how the United Daughters of the Confederacy brainwashed the youth by pretending slavery wasn't the cause of the Civil War

((2)) talks about the reason why each peek of confederate statues were created

((3)) also talks about each peek and why these statues were created

```{r}
all_dates = read.csv("data/confederate_statue_dates.csv") %>% as_tibble()

# find mode
max_statues = all_dates %>% group_by(value3) %>%
  count %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  filter(n == max(n))

mode = max_statues %>% 
  pull(value3)

with_max = max_statues %>% 
  pull(n)
```

Also, we witness the most statues being built in a single year in `r mode` with `r with_max` statues.

These bins are of width = 5 years.

```{r graph1, fig.width=20, fig.height=15}
all_dates %>%
  ggplot(aes(x = value3)) +
  geom_histogram(binwidth = 5, fill = "red") +
  scale_x_continuous(breaks = seq(min(all_dates$value3), max(all_dates$value3), by = 5)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size=25, angle = 90),
        axis.text.y = element_text(size=30),
        title = element_text(size=30)) +
  ylab("# of Statues") +
  ggtitle("Number of Confederate Statues Over Time")
```

It's just something to think about...

Also web scraping is fun



sources:  
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Alabama
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Georgia
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Mississippi
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_North_Carolina
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_South_Carolina
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials

